### 2022-03-16 satfol/satfol.tex
hle

\newcommand{\eqlit}[2]{\ensuremath{#1 \eq #2}}
\newcommand{\neqlit}[2]{\ensuremath{#1 \noteq #2}}
% \newcommand{\eqneqlit}[2]{\ensuremath{#1 \mathrel{\dot{\eq}} #2}}
\newcommand{\pospredlit}[1]{{#1}}
\renewcommand{\neglit}[1]{\neg{\kern.2ex #1}}
\newcommand{\negpredlit}[1]{\neglit{#1}}
\newcommand{\arbpredlit}[1]{(\neg)\kern.2ex\pospredlit{#1}}
\newcommand{\cor}{\ensuremath{\mathrel{\lor}}}
\newcolumntype{d}{D{.}{.}{0}}

@@

As we are working with clausal logic, we assume the clausal structure as defined
in Sect.~\ref{sec:pre:clauses}. Recall that predicate literals  are encoded
as (dis)equations with equality. To simplify the notation in this chapter we write
predicate literals in unencoded form: positive literal $\eqlit{s}{\itrue}$ is written
as $s$ and negative literal $\neqlit{s}{\top}$ is written as $\negpredlit{s}$.
Literals $\eqlit{s}{t}$ where neither $s$ nor $t$ have the type $o$ are called \emph{functional}.
Given a literal $L$, we overload notation and write $\neglit{L}$ to
denote its complement. Clauses are often defined
as sets of literals, but superposition needs multisets; with multisets,
an instance $\sigmacl{C}$ always has the same number of literals as $C$, a most
convenient property. Given a clause set $N$, $\binset{N}$ denotes the subset of
its binary clauses: $\binset{N} = \{ L_1 \cor L_2 \mid L_1 \cor L_2 \in N \}$.

@@

% \ourpara{Superposition Provers}
%
% Superposition \cite{bg-94-superposition} is a calculus for clausal
% first-order logic that extends ordered resolution
% \cite{bg-01-resolution} with equality reasoning. It is
% refutationally complete: Given a finite, unsatisfiable clause set, it will
% eventually derive the empty clause. It is parameterized by a \emph{selection
% function} that influences which of a clause's literals are eligible as the target
% of inferences. Moreover, it is compatible with the \emph{standard redundancy
% criterion}, which can be used to delete a clause $C$ while preserving completeness of the calculus.

% \looseness=-1
% The redundancy criterion relies on an order $\succ$ that
% compares terms, literals, or clauses.
% The order is used to determine whether clauses can be deleted.
% If $N$ is ground, $C$ can be deleted
% if it is entailed by $\prec$-smaller clauses in $N.$ This definition is lifted
% to nonground sets $N$. The criterion can be used to delete a
% clause that is \emph{subsumed} by another clause (e.g.,
% $\cst{\cst{p}(\cst{a})} \cor \cst{q}$ by $\cst{p}(x)$) or
% to \emph{simplify} a clause $C$ into $C'$, which amounts to adding $C'$ and
% then deleting $C$ as redundant with respect to $N \cup \{C'\}$.
% Subsumption and simplification are the main inprocessing mechanisms available
% to superposition provers. Some provers also implement clause splitting
% \cite{riazanov-voronkov-2001,fietzke-weidenbach-2009,av-2014-avatar}.

% Superposition provers saturate the input problem with respect to the calculus's
% inference rules using the \emph{given clause procedure}
% \cite{mcw-1997-otter,adf-1995-discount}. It partitions the proof state into a
% passive set $\mathcalx{P}$ and an active set $\mathcalx{A}$. All clauses start in
% $\mathcalx{P}$. At each iteration of the procedure's main loop, the prover
% chooses a clause $C$ from $\mathcalx{P}$, simplifies it, and moves it to
% $\mathcalx{A}$. Then all inferences between $C$ and active clauses are performed.
% The resulting clauses are again simplified and put in $\mathcalx{P}$.\confrep{}{
% The provers differ in which clauses are used for simplification: Otter-loop
% \cite{mcw-1997-otter} provers use both active and passive clauses whereas
% DISCOUNT-loop \cite{adf-1995-discount} provers use only active clauses.}

@@

>>>>

\newcommand{\eqlit}[2]{\ensuremath{#1 \eq #2}}
\newcommand{\neqlit}[2]{\ensuremath{#1 \noteq #2}}
% \newcommand{\eqneqlit}[2]{\ensuremath{#1 \mathrel{\dot{\eq}} #2}}
\newcommand{\pospredlit}[1]{{#1}}
\renewcommand{\neglit}[1]{\neg{\kern.2ex #1}}
\newcommand{\negpredlit}[1]{\neglit{#1}}
\newcommand{\arbpredlit}[1]{(\neg)\kern.2ex\pospredlit{#1}}
\newcommand{\cor}{\ensuremath{\mathrel{\lor}}}
\newcolumntype{d}{D{.}{.}{0}}
\definecolor{light-gray}{gray}{0.925}
\newcommand\MAX[1]{\smash{\setlength{\fboxsep}{.3ex}\colorbox{light-gray}{\ensuremath{\vphantom{('q}{#1}}}}}

@@

As we are working with clausal logic, we assume the clausal structure as defined
in Sect.~\ref{sec:pre:clauses}. In this section we use the uppercase letter $L$
to denote literals, as this is more customary in the related literature for
propositional logic. Recall that predicate literals  are encoded as
(dis)equations with equality. To simplify the notation in this chapter we write
predicate literals in unencoded form: positive literal $\eqlit{s}{\itrue}$ is
written as $s$ and negative literal $\neqlit{s}{\top}$ is written as
$\negpredlit{s}$. Literals $\eqlit{s}{t}$ where neither $s$ nor $t$ have the
type $o$ are called \emph{functional}. Given a literal $L$, we overload notation
and write $\neglit{L}$ to denote its complement. Clauses are often defined as
sets of literals, but superposition needs multisets; with multisets, an instance
$\sigmacl{C}$ always has the same number of literals as $C$, a most convenient
property. Given a clause set $N$, $\binset{N}$ denotes the subset of its binary
clauses: $\binset{N} = \{ L_1 \cor L_2 \mid L_1 \cor L_2 \in N \}$.

@@

@@

\section{Hidden-Literal-Based Elimination}
\label{sec:satfol:hidden-literal-based-elimination}

In propositional logic, binary clauses from a clause set $N$ can be used to
efficiently discover literals $L, L'$ for which the implication $L' \impl L$
is entailed by $N$'s binary clauses---i.e., $\binset{N} \models L' \impl L$.
Heule et al.\ \cite{hjb-2011-big-simplification} introduced the concept of \emph{hidden
literals} to capture such implications.

\begin{defi}
   Given a propositional literal $L$ and a propositional clause set~$N$, the
   set of \emph{propositional hidden literals} for $L$ and~$N$ is
   $\PHL(L, N) = \{ L' \mid L' \bigpimpl{N}^\ast L \} \setminus \{ L \}$, where $\bigpimpl{N}$ is
   defined such that $\neglit{L_1} \bigpimpl{N} L_2$ whenever $L_1 \cor L_2 \in N.$
   Moreover, $\PHL(L_1 \cor \cdots \cor L_n, N) = \bigcup_{i=1}^n
   \PHL(L_i, N)$.
\end{defi}

Heule et al.\ used a definition based on a fixpoint computation, but our
definition based on the reflexive transitive closure $\bigpimpl{N}^\ast$ of $\bigpimpl{N}$ is equivalent. Intuitively,
a hidden literal can be added to or removed from a clause without affecting its
semantics in models of~$N.$ By eliminating hidden literals from $C$, we simplify
it. By adding hidden literals to $C$, we might get a tautology $C'$ (i.e., a
valid clause:\ $\models C'$), meaning that $\binset{N} \models C$, thereby
enabling us to delete $C$. Note that $\PHL(L,N)$ is finite for a finite~$N.$

\begin{defi}
   \label{def:hle}\label{def:hte}
   Given $L' \cor L \cor C \in N,$ if $L' \in
   \PHL(L, N)$, \emph{hidden literal elimination} (HLE) replaces $N$ by $(N
   \setminus \{L' \cor L \cor C\}) \cup \{L \cor C\}$.
   Given $C \in N$, $\{L_1,\dots,L_n\} = \PHL(C, N)$, and $C' = C \cor L_1 \cor
   \cdots \cor L_n$, if $C'$ is a tautology, \emph{hidden
   tautology elimination} (HTE) replaces $N$ by $N \setminus \{C\}$.
\end{defi}

\begin{theorem}
   The result of applying HLE or HTE to a clause set $N$ is
   equivalent to $N$.
\end{theorem}

\begin{proof}
For HLE, if $L' \in \PHL(L, N)$,  $\binset{N} \models \neglit{L'} \cor L$.
Then, subsumption resolution \cite{bg-01-resolution} yields shortened clause $L \cor C'$ from
Definition~\ref{def:hle}. For HTE, it can be shown that $N' \models C$ if and
only if $C \cor L'$, where $L' \in \PHL(C, N)$. By transitivity of equivalence,
we get the desired result.
\end{proof}

We generalize hidden literals to first-order logic with equality by
considering substitutivity of variables as well as congruence of equality.

\begin{defi}
\label{def:hl-fo}
Given a literal $L$ and a clause set $N$,
the set of \emph{hidden literals} for $L$ and $N$
is $\HL(L,N) = \{ L' \mid L' \bigfimpl{N}^\ast L \} \setminus \{ L \}$, where
$\bigfimpl{N}$ is defined so that:
\begin{enumerate}
    \item $\sigmaterm{\neglit{L'}} \bigfimpl{N} \sigmaterm{L}$ if
    $L' \cor L \in N$ and $\sigma$ is a substitution
    \item $\eqlit{s}{t} \bigfimpl{N} \eqlit{u[s]}{u[t]}$ for all terms $s, t$ and
    contexts $u[\phantom{\cdot}]$
    \item $\neqlit{u[s]}{u[t]} \bigfimpl{N} \neqlit{s}{t}$ for all terms $s, t$ and
    contexts $u[\phantom{\cdot}]$
\end{enumerate}
% (1)~;
% (2)~; and
% (3)~.
Moreover, $\HL(L_1 \cor \cdots \cor L_n, N) = \bigcup_{i=1}^n \HL(L_i, N)$.
\end{defi}

\looseness=-1
The generalized definition also enjoys the key property that $L'
\in \HL(L, N)$ implies $\binset{N} \models L' \impl L$. However, $\HL(L, N)$
may be infinite even for predicate literals; for example,
$\pospredlit{\cst{p}(\cst{f}^i(x))} \in \HL(\pospredlit{\cst{p}(x)}, \{\pospredlit{\cst{p}(x)} \cor
\negpredlit{\cst{p}(\cst{f}(x))} \})$ for every $i$.

Based on Definition \ref{def:hl-fo}, we can generalize hidden literal elimination
and support a related technique:
\begin{align*}
    & \namedsimpsc{HLE}{L' \cor L \cor C}{L \cor C}{L' \in \HL(L,N)} \\[.5\jot]
    & \namedsimpsc{\rlap{FLE}\phantom{HLE}}{L \cor C}{C}{L',\neglit{L'} \in \HL(\neglit{L}, N)}
 \end{align*}
Recall that double lines denote simplification rules (Sect.~\ref{sec:pre:sup}).
The second rule is called \textit{failed literal elimination},
inspired by the SAT technique of asserting $\neglit{L}$ if $L$ is a \textit{failed literal} \cite{jwf-1995-fld}.
It is easy to see that rule \infname{HLE} is sound.
From $L' \in \HL(L,N)$ we have $N \models L' \impl L$ (i.e., $\neglit{L'} \cor L$).
Performing subsumption resolution \cite{bg-01-resolution} between $L'
\cor L \cor C$ and $\neglit{L'} \cor L$ yields the conclusion, which is
therefore entailed by $N.$ For \infname{FLE}, the condition $L',
\neglit{L'} \in \HL(\neglit{L},N)$ means that $\binset{N} \models \{\neglit{L'} \cor
\neglit{L}{,}\; L' \cor \neglit{L}\} \models \neglit{L}$.

\begin{exa}
   \label{example:hle}
   Consider the clause set $N = \{ \cst{p}(x) \cor \neglit{\cst{p}(\cst{f}(x))}{,}\;
   \cst{p}(\cst{f}(\cst{f}(x))) \cor \eqlit{\cst{a}}{\cst{b}} \}$ and the clause
   $C = \neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})} \cor \cst{p}(x)$. The first
   clause in $N$ induces $\cst{p}(\cst{f}(x)) \bigfimpl{N} {\cst{p}(x)}$,
   $\cst{p}(\cst{f}(\cst{f}(x))) \bigfimpl{N} \cst{p}(\cst{f}(x))$, and
   hence $\cst{p}(\cst{f}(\cst{f}(x))) \bigfimpl{N}^\ast  \cst{p}(x)$.
   Together with the second clause in $N$, it can be used to derive $\neqlit{\cst{a}}{\cst{b}}
   \bigfimpl{N}^\ast  \cst{p}(x)$. Finally, using rule~(3) of Definition
   \ref{def:hl-fo}, we derive $\neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})}
   \bigfimpl{N}^\ast  \cst{p}(x)$---that is,
   $\neqlit{\cst{f}(\cst{a})}{\cst{f}(\cst{b})} \in \HL(\cst{p}(x), N)$. This
   allows us to remove $C$'s first literal using \infname{HLE}.
\end{exa}

Two special cases of \infname{HLE} exploit equality congruence as embodied by
conditions 2 and 3 of Definition~\ref{def:hl-fo} without requiring to
compute the $\HL$~set:
\begin{align*}
  %% TYPESETTING: \,'s below
  \namedsimp{CongHLE$^+$}{\eqlit{s}{t} \cor \eqlit{u[s]}{u[t]} \cor C}{\eqlit{u[s]}{u[t]} \cor C}\,\, \\[.5\jot]
  \namedsimp{CongHLE$^-$}{\neqlit{s}{t} \cor \neqlit{u[s]}{u[t]} \cor C}{\neqlit{s}{t} \cor C}
\end{align*}%

Hidden literals can be combined with unit clauses~$L'$ to remove more literals:
\[\namedsimpsc{UnitHLE}{L' \quad L \cor C}{L' \quad C}{\sigmaterm{L'} \in \HL(\neglit{L},N)}
%\text{~for some~}\sigma -- %%% Not necessary in a rule -- all variables are universally quantified. --JB
\]
%
Given a unit clause $L' \in N$, the rule uses it to discharge $\sigmaterm{L'}$ in
$N \models \sigmaterm{L'} \impl \neglit{L}$. As a result, we have $N \models
\neglit{L}$, making it possible to remove $L$ from $L \cor C$.

% Kiesl and Suda described first-order generalization of {\it asymmetric literals}
% \cite{ks-2017-unif-principle}: literal $L\sigma$ is asymmetric for a clause $C$
% and a clause set $N$ if there exists a clause $D \cor \neglit{L} \in N \setminus
% \{C \}$ such that $D\sigma \subseteq C$ for some substitution $\sigma$.

\begin{exa}
   \label{example:unithle}
   \looseness=-1
   Consider the clause set $N = \{\cst{p}(x) \cor \cst{q}(\cst{f}(x)){,}\;
   \neglit{\cst{q}(\cst{f}(\cst{a}))} \cor \eqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})}{,}\;
   \neqlit{\cst{f}(x)}{\cst{g}(y)}\}$ and the clause $C = \neglit{\cst{p}(\cst{a})} \cor
   \neglit{\cst{q}(\cst{b})}$. The first clause in $N$ induces
   $\neglit{\cst{q}(\cst{f}(\cst{a}))} \bigfimpl{N} \cst{p}(\cst{a})$, whereas
   the second one induces $\neqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})} \bigfimpl{N}
   \neglit{\cst{q}(\cst{f}(\cst{a}))}$. Thus, we have $\neqlit{\cst{f}(\cst{b})}{\cst{g}(\cst{c})}
   \bigfimpl{N}^\ast \cst{p}(\cst{a})$---that is, $\neqlit{\cst{f}(\cst{b})}{\cst{f}(\cst{c})} \in
   \HL(\cst{p}(\cst{a}), N)$. By applying the substitution $\{x \mapsto \cst{b}{,}\allowbreak\; y
   \mapsto \cst{c}\}$ to the third clause in $N$, we can fulfill the conditions of
   \infname{UnitHLE} and remove $C$'s first literal.
\end{exa}

Next, we generalize hidden tautologies to first-order logic.

\begin{defi}
   A clause $C$ is a \emph{hidden tautology} for a clause set $N$ if there
   exists a finite set $\{L_1,\dots,L_n\} \subseteq \HL(C,N)$ such that
   $C \cor L_1 \cor \cdots \cor L_n$ is a tautology.
\end{defi}

\begin{exa}
   \label{example:hidden-tautologies-completeness}
   \looseness=-1
   In general, hidden tautologies are not redundant and cannot be deleted during saturation.
   Consider the unsatisfiable set $N = \{ \negpredlit{\cst{a}}{,}\;
   \negpredlit{\cst{b}}{,}\; \pospredlit{\cst{a}} \cor
   \pospredlit{\cst{c}}{,}\allowbreak\; \pospredlit{\cst{b}} \cor \negpredlit{\cst{c}}
   \}$, the order $\cst{a} \prec
   \cst{b} \prec \cst{c}$, and the empty selection function. The
   only possible superposition inference from $N$ is between the last two clauses,
   yielding the hidden tautology $\pospredlit{\cst{a}} \cor \pospredlit{\cst{b}}$
   (after simplifying away $\neqlit{\top}{\top}$), which is entailed by the larger
   clauses $\pospredlit{\cst{a}} \cor \pospredlit{\cst{c}}$ and
   $\pospredlit{\cst{b}} \cor \negpredlit{\cst{c}}$. If this clause is removed,
   the prover could enter an infinite loop, forever generating and deleting the
   hidden tautology\confrep{}{ and never getting the opportunity to derive the
   empty clause}.
\end{exa}

\begin{rep}
   \newcommand{\aset}{\mathcal{A}}
   \newcommand{\pset}{\mathcal{P}}

   In practice, most provers use a variant of the given clause procedure.
   Removing hidden tautologies breaks the invariant of the procedure that all
   inferences between clauses in $\aset$ are redundant. The end result is not
   that the prover diverges, but that it terminates without deriving empty
   clause.

   To observe this, assume the setting as in Example
   \ref{example:hidden-tautologies-completeness}, and let $\pset=N$ and
   $\aset=\emptyset$. After moving the first three clauses from $\pset$ to
   $\aset$ ($\aset = \{ \negpredlit{\cst{a}}, \negpredlit{\cst{b}}, \pospredlit{\cst{a}} \cor
   \pospredlit{\cst{c}} \}$, $\pset = \{\pospredlit{\cst{b}} \cor \negpredlit{\cst{c}}\}$), 
   no inferences are possible, and no new clauses are added to $\pset$. After the last clause
   is moved to $\aset$, the hidden tautology $\pospredlit{\cst{a}} \cor \pospredlit{\cst{b}}$
   is produced. If it is deleted, the prover terminates with the unsatisfiable set $\aset$,
   but does not derive the empty clause. 
\end{rep}

To delete hidden tautologies during saturation, the prover could check that all
the relevant clause instances encountered along the computation of $\HL$ are
$\prec$-smaller than a given hidden tautology. However, this would be expensive
and seldom succeed, given that superposition creates lots of nonredundant
hidden tautologies. Instead, we propose to simplify hidden tautologies using the
following rules:
%
%\begin{linenomath*}
% \begin{align*}
%    %% TYPESETTING: \,'s below
%    \namedsimpsc{HTR}{L \cor L' \cor C}{L \cor L'}{\neglit{L'} \in \HL(L, N)} &\\
%    \namedsimpsc{FLR}{L \cor C}{L}{L',\neglit{L'} \in \HL(L, N)} &
% \end{align*}
%
\begin{align*}
   & \namedsimpsc{HTR}{L \cor L' \cor C}{L \cor L'}{\neglit{L'} \in \HL(L, N) \text{~and~} C \not= \bot} \\[.5\jot]
   & \namedsimpsc{\rlap{FLR}\phantom{HTR}}{L \cor C}{L}{L',\neglit{L'} \in \HL(L, N) \text{~and~} C \not= \bot}
\end{align*}
%

%\end{linenomath*}
%
We call these techniques \emph{hidden tautology reduction} and \emph{failed literal
reduction}, respectively.
% \infname{FLD} is inspired by the \confrep{}{eponymous }SAT
% approach of asserting $L$ if $\neglit{L}$ is determined to be a failed literal
% \cite{jwf-1995-fld}.
%
Both rules are sound.
%
As with hidden literals, unit clauses~$L'$ can be exploited:
\[\namedsimpsc{UnitHTR}{L' \quad L \cor C}{L' \quad L}{\sigmaterm{L'} \in \HL(L,N) \text{~and~} C \not= \bot}
%\text{~for some~}\sigma
\]

We give the simplification rules above\confrep{}{ (for hidden literal
elimination, hidden tautology reduction,
failed literal detection, and their variants)} the collective name of
\emph{hidden-literal-based elimination} (HLBE).
Yet another use of hidden literals is for \emph{equivalent literal
substitution} \cite{hjb-2010-cl-elim}: If both $L' \in \HL(L, N)$ and $L
\in \HL(L', N)$, we can often simplify $\sigmacl{L'}$ to $\sigmacl{L}$ in $N$ if
$\sigmacl{L'} \succ \sigmacl{L}$. We want to investigate this~further.

% \looseness=-1
% Provers based on the DISCOUNT loop use only active clauses for
% simplification. If we restrict our attention to a finite subset of hidden
% literals, we can organize them into data structures that support efficient
% lookup of a hidden literal. Efficient implementation of these data structures
% allows to use all passive clauses for simplification, strengthening DISCOUNT
% loop simplification machinery.

\begin{theorem}
\begin{sloppypar}
The rules \infname{HLE}, \infname{FLE}, \infname{CongHLE$^+$},
\infname{Cong\allowbreak HLE$^-$}, \infname{UnitHLE}, \infname{HTR},
\infname{FLR}, and \infname{UnitHTR} are sound simplification rules.
\end{sloppypar}
\end{theorem}

\begin{rep}
\begin{proof}
It is easy to see that the deleted premises are entailed by the conclusions that
replace them and that the conclusions' instances are $\prec$-smaller than the
premises' instances, as required by the redundancy criterion. It remains to
check soundness.

\medskip

\noindent
\textsc{Case} \infname{HLE}:\enskip
We have $N{,}\> L' \models L$ by the side condition and must show
$N{,}\> L' \lor L \lor C \models L \lor C$. Let
$\ourmodel \models N{,}\> L' \lor L \lor C$. If $\ourmodel \models L'$, then
we also have $\ourmodel \models L$ thanks to the side condition and hence
$\ourmodel \models L \lor C$. Otherwise, we have $\ourmodel \models L \lor C$,
which is exactly what we need to show.

\medskip

\noindent
\textsc{Case} \infname{FLE}:\enskip
We have $N{,}\> L \models L'$ and $N, L \models \lnot L'$ by the side condition.
If $\ourmodel \models N{,}\> L$, then both $\ourmodel \models L'$ and
$\ourmodel \models \lnot L'$, an absurdity.
Otherwise, we have $\ourmodel \models C$, as desired.

\medskip

\noindent
\textsc{Case} \infname{CongHLE$^+$}, \infname{CongHLE$^-$}:\enskip
Obvious by congruence of equality.

\medskip

\noindent
\textsc{Case} \infname{UnitHLE}:\enskip
We have $N{,}\> L \models \lnot \sigmaterm{L'}$ by the side condition. If $\ourmodel
\models N{,}\> L$, then $N \models \lnot \sigmaterm{L'}$. But since $L' \in N$, this is
an absurdity. Otherwise, we have $\ourmodel \models C$, as desired.

\medskip

\noindent
\textsc{Case} \infname{HTR}:\enskip
We have $N{,}\> \lnot L' \models L$ by the side condition. If either $\ourmodel
\models L$ or $\ourmodel \models L'$, the desired result follows directly.
Otherwise, from $\ourmodel \models \lnot L'$ we also have $\ourmodel \models L$
thanks to the side condition, contradicting $\ourmodel \models \lnot L$.

\medskip

\noindent
\textsc{Case} \infname{FLR}:\enskip
We have $N{,}\> L' \models L$ and $N{,}\> \lnot L' \models L$ by the side condition.
Hence $N \models L$, as desired.

\medskip

\noindent
\textsc{Case} \infname{UnitHTR}:\enskip
We have $N{,}\> \sigmaterm{L'} \models L$. Since $L' \in N$, we have $N \models L$,
as desired.
\qedhere
\end{proof}
\end{rep}
